Efficient GPU implementation of multicomponent and multiphase Lattice Boltzmann models.

### 2010-04-28 paper.tex
Minor typographical fixes.

The lattice Boltzmann equation operates on the rectangular grid representing the physical domain. It utilizes the probability distribution functions containing the information about the macroscopic variables, such as density, momentum. LBE consists from two parts: local collision and propagation operated from one node to another in the certain direction specified by the velocity set.  The lattice Boltzmann implementation in a standard view representing the collision and propagation steps:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) + F_i,&&\text{ collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $F_i$ is the external force population. The equilibrium function population is usually given as \cite{}:
\begin{equation}
f_i^{eq}=w_i \rho \biggl(1+\frac{c_{i\alpha}u_{\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\biggr),
\end{equation}
where for D2Q9 $w_i=\{\dfrac{4}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36}\}$ and $Q_{i\alpha\beta}=$ with the sound speed parameter $c_s^2=\dfrac{1}{3}$. The macroscopic variables are restored on every time step through the probability function population:
\begin{equation*}
\begin{aligned}
&\rho=\sum_i{f_i}
&\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}
\end{aligned}
\end{equation*}
The multiphase models can be simulated either through the force population and/or special equilibrium function. The Shan-Chen model utilizes the force population, but the free-energy model is simulated through the special equilibrium model. Both models are described in the next sections.

\subsection{Shan-Chen model}
In the Shan-Chen model \cite{Shan-chen:extended} the force at a given node depends
on all local neighbours characteristics:
\begin{equation}  \label{Shan-Chen:Shan-Chen:cont}
\bm{F}(\bm{x})=-G\psi(\rho(\bm{x}))\sum_{i}{w_i \psi(\rho(\bm{x}+\bm{c_i}))\bm{c_{i}}},
\end{equation}
where $\psi$ is the function of local characteristics of a node, and taken
as $\psi(\rho)=1-\exp(-\rho)$. Coefficient $G$ is responsible for the attraction
force between the molecules. Overall, the Shan-Chen force is the approximation of
the force between molecules \cite{kwok,kwok-contact-angle}. The force modifies
the momentum fluxes as:
\begin{equation*}
P_{\alpha\beta}=\biggl(c_s^2\rho+\frac{G}{6}\psi^2+\frac{G}{36}%
|\bm{\nabla}\psi|^2 +\frac{G}{18}\psi\Delta\psi\biggr)\delta_{\alpha\beta}-\frac{G}{%
18}\partial_{\alpha}\psi\partial_{\beta}\psi,
\end{equation*}
where the equation of states is changed from original $P=c_s^2 \rho$ \cite{Succi-book}
to the van-der-Waals type equation of state:
\begin{equation}
P=c_s^2 \rho +\frac{G}{6} \psi(\rho)^2.
\end{equation}
The lattice Boltzmann simulation for the Shan-Chen model is implemented through the standard implementation (\ref{standard:implementation}). In the presence of the external force the macroscopic velocity is changed as $\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}+\dfrac{F_{\alpha}}{2}$. The force population is defined as \cite{guo}:
\begin{equation}
F_i=w_i\biggl(1-\frac{\omega}{2}\biggr)\biggl(3 (c_{i\alpha}-u_{\alpha}) + 9 (c_{i\beta}u_{\beta})c_{i\alpha}\biggr)F_{\alpha}.
\end{equation}
The Shan-Chen model is one-component multiphase model. The multicomponent modification is presented in the Appendix~\ref{sec:binary_shan_chen}.
\subsection{Free-energy model}
The free-energy or pressure model \cite{swift} allows the incorporation of the any equation of state to the lattice Boltzmann equation. In comparison with the Shan-Chen model, the free-energy model is simulated through the special choice of the equilibrium distribution function. The lattice Boltzmann scheme is implemented in a standard form (\ref{standard:implementation}), but the equilibrium function is defined as:
\begin{equation}
\begin{aligned}
f_i^{eq}=&w_i \biggl(p_0(\rho)-3 k\rho\Delta\rho+\rho \frac{c_{i\alpha}u_{\alpha}}{3}+\rho \frac{9 Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2}\biggr)\\
&+k (w_{ixx} \partial_x \rho \partial_x \rho + w_{ixy} \partial_x \rho \partial_y \rho + w_{iyy} \partial_y \rho \partial_y \rho),
\end{aligned}
\end{equation}
where $w_{ixx}=\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\}$, $w_{iyy}=\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\}$ and $w_{ixy}=\{0,0,0,0,\dfrac{1}{4},\dfrac{1}{4},-\dfrac{1}{4},-\dfrac{1}{4}\}$. The laplacian and the gradient are taken by the numerical stencils:
\begin{equation}
\Delta=\frac{1}{6}
\begin{pmatrix}
1&4&1\\
4&-20&4\\
1&4&1\\
\end{pmatrix}\text{ and }
\partial_x=\frac{1}{12}
\begin{pmatrix}
-1&0&1\\
-4&0&4\\
-1&0&1\\
\end{pmatrix}.
\end{equation}

>>>>

The lattice Boltzmann equation operates on the rectangular grid representing the physical domain. It utilizes the
probability distribution functions containing the information about the macroscopic variables, such as density,
momentum. LBE consists from two parts: local collision and propagation operated from one node to another in the certain
direction specified by the velocity set.  The lattice Boltzmann implementation in a standard view representing the
collision and propagation steps:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) + F_i,&&\text{ collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{ propagation step}, 
\end{aligned}
\end{equation}
where $F_i$ is the external force population. The equilibrium function population is usually given as \cite{}:
\begin{equation}
f_i^{eq}=w_i \rho \biggl(1+\frac{c_{i\alpha}u_{\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\biggr),
\end{equation}
where for D2Q9
$w_i=\left\{\dfrac{4}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{9},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36},\dfrac{1}{36}\right\}$
and $Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with the sound speed parameter $c_s^2=\dfrac{1}{3}$. The macroscopic variables are restored on
every time step through the probability function population:
\begin{equation*}
\begin{aligned}
&\rho=\sum_i{f_i}
&\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}
\end{aligned}
\end{equation*}
The multiphase models can be simulated either through the force population and/or special equilibrium function. The
Shan-Chen model utilizes the force population, but the free-energy model is simulated through the special equilibrium
model. Both models are described in the next sections.

\subsection{Shan-Chen model}
In the Shan-Chen model \cite{Shan-chen:extended} the force at a given node depends
on all local neighbours characteristics:
\begin{equation}  \label{Shan-Chen:Shan-Chen:cont}
\bm{F}(\bm{x})=-G\psi(\rho(\bm{x}))\sum_{i}{w_i \psi(\rho(\bm{x}+\bm{c_i}))\bm{c_{i}}},
\end{equation}
where $\psi$ is the function of local characteristics of a node, and taken
as $\psi(\rho)=1-\exp(-\rho)$. Coefficient $G$ is responsible for the attraction
force between the molecules. Overall, the Shan-Chen force is the approximation of
the force between molecules \cite{kwok,kwok-contact-angle}. The force modifies
the momentum fluxes as:
\begin{equation*}
P_{\alpha\beta}=\biggl(c_s^2\rho+\frac{G}{6}\psi^2+\frac{G}{36}%
|\bm{\nabla}\psi|^2 +\frac{G}{18}\psi\Delta\psi\biggr)\delta_{\alpha\beta}-\frac{G}{%
18}\partial_{\alpha}\psi\partial_{\beta}\psi,
\end{equation*}
where the equation of states is changed from original $P=c_s^2 \rho$ \cite{Succi-book}
to the van-der-Waals type equation of state:
\begin{equation}
P=c_s^2 \rho +\frac{G}{6} \psi(\rho)^2.
\end{equation}
The lattice Boltzmann simulation for the Shan-Chen model is implemented through the standard implementation (\ref{standard:implementation}). In the presence of the external force the macroscopic velocity is changed as $\rho u_{\alpha}=\sum_i{f_i c_{i\alpha}}+\dfrac{F_{\alpha}}{2}$. The force population is defined as \cite{guo}:
\begin{equation}
F_i=w_i\biggl(1-\frac{\omega}{2}\biggr)\biggl(3 (c_{i\alpha}-u_{\alpha}) + 9 (c_{i\beta}u_{\beta})c_{i\alpha}\biggr)F_{\alpha}.
\end{equation}
The Shan-Chen model is one-component multiphase model. The multicomponent modification is presented in the Appendix~\ref{sec:binary_shan_chen}.
\subsection{Free-energy model}
The free-energy or pressure model \cite{swift} allows the incorporation of the any equation of state to the lattice Boltzmann equation. In comparison with the Shan-Chen model, the free-energy model is simulated through the special choice of the equilibrium distribution function. The lattice Boltzmann scheme is implemented in a standard form (\ref{standard:implementation}), but the equilibrium function is defined as:
\begin{equation}
\begin{aligned}
f_i^{eq}=&w_i \biggl(p_0(\rho)-3 k\rho\Delta\rho+\rho \frac{c_{i\alpha}u_{\alpha}}{3}+\rho \frac{9 Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2}\biggr)\\
&+k (w_{ixx} \partial_x \rho \partial_x \rho + w_{ixy} \partial_x \rho \partial_y \rho + w_{iyy} \partial_y \rho \partial_y \rho),
\end{aligned}
\end{equation}
where $w_{ixx}=\left\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$,
$w_{iyy}=\left\{\dfrac{1}{3},\dfrac{1}{3},-\dfrac{1}{6},-\dfrac{1}{6},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24},-\dfrac{1}{24}\right\}$
and $w_{ixy}=\left\{0,0,0,0,\dfrac{1}{4},\dfrac{1}{4},-\dfrac{1}{4},-\dfrac{1}{4}\right\}$. The laplacian and the gradient are taken by the numerical stencils:
\begin{equation}
\Delta=\frac{1}{6}
\begin{pmatrix}
1&4&1\\
4&-20&4\\
1&4&1\\
\end{pmatrix}\text{ and }
\partial_x=\frac{1}{12}
\begin{pmatrix}
-1&0&1\\
-4&0&4\\
-1&0&1\\
\end{pmatrix}.
\end{equation}
